### 2017-04-03 paper.tex
Add basic packet stuf

@@

\begin{enumerate}
\item provide a system to transmit binary data.
\item minimise the amount of overhead and bandwidth required.
\item allow multiple payload's data to be sent on the same link.
\item provide stable transmissions on a low-power link.
\item correct as many transmission-induced data errors as possible.
\item allow ground crew to track the balloon's location over time.
\end{enumerate}

@@

% TODO: references for DominoEx and Olivia

@@

\subsubsection{Frame Layer}

@@

% TODO: explain requirement for sync marker

@@

This preliminary design proved impossible to work with. While it encapsulates
the required data and metadata, it does not provide any way to identify the
start or the end of a frame in a raw binary stream, which would have rendered
parsing an incoming frame stream impossible.

@@

\begin{figure}[H]
    \begin{center}
    \begin{bytefield}[bitwidth=1.8em]{20}
        \bitbox{8}{frame} &
        \bitbox{8}{frame} &
        \bitbox{4}{...}
    \end{bytefield}
    \end{center}
    \centering
    \caption{Frame stream with embedded start sequence}
    \label{fig:frame-sync-emb}
\end{figure}

@@

\begin{figure}[H]
    \begin{center}
    \begin{bytefield}[bitwidth=1.8em]{20}
        \bitbox{1}{\texttt{AA}} &
        \bitbox{8}{frame} &
        \bitbox{1}{\texttt{AA}} &
        \bitbox{8}{frame} &
        \bitbox{2}{...}
    \end{bytefield}
    \end{center}
    \centering
    \caption{Frame stream with external synchronisation byte}
    \label{fig:frame-sync-extra}
\end{figure}

@@

\begin{figure}[H]
    \begin{bytefield}{16}
        \bitheader{0,7,8,15} \\
        \begin{rightwordgroup}{Header}
            \bitbox{8}{\texttt{0x5A}} & \bitbox{8}{version} \\
            \bitbox{16}{sequence number}
        \end{rightwordgroup} \\
        \begin{rightwordgroup}{Payload}
            \wordbox[lrt]{1}{data} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{rightwordgroup} \\
        \begin{rightwordgroup}{FEC}
            \wordbox[lrt]{1}{checksum} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{rightwordgroup}
    \end{bytefield}
    \centering
    \caption{AHABus Frame Format - Version 3}
    \label{fig:frame-fmt-3}
\end{figure}

@@

% Topic list for this:
%   * Provide reliability:
%       * Data arrives whole
%       * We receive valid data
%   * receive all data: we chunk the data stream, number the chunks         v
%   * we tag data with housekeeping things                                  v~
%   * If frames too small, housekeeping/data ratio is shit
%   * Need a way to detect the start + end of a frame.
%   * Explain that end of frame marker complicated, going with count?
%
%   * Validity checked (and corrected using FEC)
%   * FEC: went with Reed-Solomon for simplicity + availability
%   * FEC means we need small blocks of data -> Frames
%
%   * FEC forces us to change from double marker inside to sync sequence + frame

@@

\begin{figure}[H]
    \begin{bytefield}{32}
        \bitheader{0,7,8,15,16,23,24,31} \\
        \begin{leftwordgroup}{Header}
            \bitbox{8}{version} & \bitbox{8}{payload ID} &
            \bitbox{16}{packet length} \\
            \bitbox{32}{latitude} \\
            \bitbox{32}{longitude} \\
            \bitbox{16}{altitude} & \bitbox[lrt]{16}{}
        \end{leftwordgroup} \\
        \begin{leftwordgroup}{Payload}
            \wordbox[lrt]{1}{data} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{leftwordgroup}
    \end{bytefield}
    \centering
    \caption{AHABus Frame Format}
\end{figure}

@@

\section{Bus and Flight Software}

@@

\section{Flight Software}

>>>>

% TODO: add something about:
%       - hardware choices
%       - software choices
%       - language choices
%
%       - overall arch (flight platform, ground segmetn)

@@

\begin{enumerate}
\item provide a system to transmit binary data.% DATA LINK
\item minimise the amount of overhead and bandwidth required.% ALL
\item allow multiple payload's data to be sent on the same link.% PACKET LAYER
\item provide stable transmissions on a low-power link.% FRAME LAYER
\item correct as many transmission-induced data errors as possible.% FRAME LAYER
\item allow ground crew to track the balloon's location over time.% PACKET LAYER
\end{enumerate}

@@

@@

\subsubsection{Frame Layer}
\label{sssec:frame-layer}

@@

@@

This preliminary design proved impossible to work with: while it encapsulates
the required data and metadata, it does not provide any way to identify the
start or the end of a frame in a raw binary stream, which would have rendered
parsing an incoming frame stream impossible.

@@

\begin{figure}[H]
    \begin{center}
    \begin{bytefield}[bitwidth=1.8em]{20}
        \bitbox{1}{...} &
        \bitbox{8}{frame} &
        \bitbox{8}{frame} &
        \bitbox{3}{...}
    \end{bytefield}
    \end{center}
    \centering
    \caption{Frame stream with embedded start sequence}
    \label{fig:frame-sync-emb}
\end{figure}

@@

\begin{figure}[H]
    \begin{center}
    \begin{bytefield}[bitwidth=1.8em]{20}
        \bitbox{1}{...} &
        \bitbox{1}{\texttt{AA}} &
        \bitbox{8}{frame} &
        \bitbox{1}{\texttt{AA}} &
        \bitbox{8}{frame} &
        \bitbox{1}{...}
    \end{bytefield}
    \end{center}
    \centering
    \caption{Frame stream with external synchronisation byte}
    \label{fig:frame-sync-extra}
\end{figure}

@@

\begin{figure}[H]
    \begin{bytefield}{16}
        \bitheader{0,7,8,15} \\
        \begin{leftwordgroup}{Header}
            \bitbox{8}{\texttt{0x5A}} & \bitbox{8}{version} \\
            \bitbox{16}{sequence number}
        \end{leftwordgroup} \\
        \begin{leftwordgroup}{Payload}
            \wordbox[lrt]{1}{data} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{leftwordgroup} \\
        \begin{leftwordgroup}{FEC}
            \wordbox[lrt]{1}{checksum} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{leftwordgroup}
    \end{bytefield}
    \centering
    \caption{AHABus Frame Format - Version 3}
    \label{fig:frame-fmt-3}
\end{figure}

%TODO: what's that worth? shouldn't it be at the end

@@

@@

The Packet Layer relies on the stable data stream provided by the Link and Frame
Layers, and was designed to allow the multiplexing of multiple payload's data as
well as the transmission of ancillary data required for the overall flight
management (platform's GNSS co-ordinates and altitude).

To achieve those goals, the AHABus protocol uses packets similar to frames. Each
packet encapsulates a chunk of data retrieved from a single payload. To that
end, packets were designed to be of variable length, to avoid unnecessary
padding, which would have wasted bandwidth. The data - referred to as the
packet's \textit{payload} – is preceded by the packet's header, which contains
information necessary to attribute the packet's data to a payload, indicate
the length of the packet, and carry the AHABus platform's location and altitude.

This design achieves multiplexing through time division: packets are chunked
and sent in frames one after the other, and can be dispatched to their
respective payload teams once decoded at the ground station. The main issue with
that method is that the protocol itself does not ensure that each payload gets
an even or fair share of the total bandwidth. As discussed in section
\ref{sec:flight-software}, this was solved in the flight software design.

\begin{figure}[H]
    \begin{bytefield}{32}
        \bitheader{0,7,8,15,16,23,24,31} \\
        \begin{leftwordgroup}{Header}
            \bitbox{8}{version} & \bitbox{8}{payload ID} &
            \bitbox{16}{length} \\
            \bitbox{32}{latitude} \\
            \bitbox{32}{longitude} \\
            \bitbox{16}{altitude} & \bitbox[lrt]{16}{}
        \end{leftwordgroup} \\
        \begin{leftwordgroup}{Payload}
            \wordbox[lrb]{3}{data[length]}
        \end{leftwordgroup}
    \end{bytefield}
    \centering
    \caption{AHABus Packet Format - Version 1}
    \label{fig:packet-fmt-1}
\end{figure}

Figure \ref{fig:packet-fmt-1} shows the first and current version of the AHABus
packet format. Like frame discussed in section \ref{sssec:frame-layer}, each
packet carries a version flag which was added to allow backwards compatibility,
were the protocol to evolve in the future.

The encoding of the metadata was chosen based on the physical constraints of
the information:

\begin{itemize}
\item Latitude and Longitudes can be both positive (North, East) or Negative
(South, West). At the equator, \({0.0001}^\circ\) \footnotemark of
longitude maps to roughly 11 metres, and that distance decreases when moving
towards the poles, which was deemed precise enough for the project's purpose.
To avoid conflicts in floating-point mathematics conventions, and allow packet
encoders to be written for embedded platform without hardware floating point
support, latitude and longitude were designed to be multiplied by \(10,000\)
and transmitted as 32-bit precision signed integers.

\item Altitude is positive for the whole mission duration, and cannot exceed
\(40,000m\). Since a sub-metre precision is not required, and would hardly be
achievable with civilian GNSS receivers, it is encoded as a 16-bit precision
unsigned integer.
\end{itemize}

\footnotetext{
Earth circumference at the equator: \(40\time10^{6}m\).
\(D_{0.0001^{\circ}} = 0.0001\frac{40\time10^{6}}{360}=11.1...m\)
}

@@

\section{Payload Data Bus}

@@

\section{Flight Software}
\label{sec:flight-software}
